Luminosity function and flux function

### 2015-12-04 luminosity.tex
minor updates
Update title & affil
text edit

\title{Luminosity function of RGHJs}
\author{%
Yuka Fujii\altaffilmark{1}, %
David S. Spiegel\altaffilmark{2,3,4}%
}

@@

\begin{abstract}
We examine how the highest observed flux evolves as we observe further. 
Given a luminosity function characterized by a power law with lower and upper limits of luminosity, firstly we consider the distribution of observed flux from a sphere within distance $R$ of a given luminosity function ("flux function"). The flux function composes of three regimes: in order of increasing flux, 1) zero flux, 2) a power law which depends on the luminosity function, and 3) a power law with index -2.5. 
Then, we compute the expected value of the highest flux from a spherical shell at distance $R\sim R+dR$. within a volume with radius $R$. 
\end{abstract}

\keywords{planetary systems --- planets and satellites: general}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Distribution of observed flux} 
\label{sec:luminosity function}

@@

In this section, we consider the number density of observed flux of the targets at spherical shell at distance $r$.

In the following, Luminosity and (observed) flux are denoted by $L$ and $F$, respectively. 
\vspace{1\baselineskip}

@@

Now, we want to change variables from $\{L,\,r\}$ to $\{F,\,r\}$. 
The relationship between flux $F$ and luminosity $L$ is, as usual,
%%%
\begin{equation}
F = \frac{L}{4\pi r^2} 
\end{equation}
%%%
where $r$ is the distance to the target. 
As changing the set of variables needs to employ Jacobian,
%%%
\begin{eqnarray}
\frac{d^2 N}{dL dr} &=& \Psi (4\pi r^2F', r)\; det(J) \\
det(J) &=& \left| \PD{r}{r}\PD{L}{F'} - \PD{r}{F'}\PD{L}{r} \right| = 4 \pi r^2
\end{eqnarray}
%%%
Therefore, the number density of the flux coming from the spherical shell is
%%%
\begin{equation}
\frac{d^2 N}{dF dr} = \left\{
\begin{array}{ll}
C' F^{-\alpha} r^{4-2\alpha } & (L_{\rm min} < F/4\pi r^2 < L_{\rm max}) \label{eq:dNdFdr} \\
0 & (\mbox{elsewhere})
\end{array}
\right. 
\end{equation}
%%%
where $C' \equiv (4 \pi )^{2-\alpha}$.

@@

%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{dndF_alpha31.pdf}
   \plotoneh{dndF_alpha15.pdf}
   \caption{Number density of observed flux in $\mu {\rm Jy}$.}
  \label{fig:dndF}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Examples with $\alpha = 3.1$ and $\alpha = 1.5$ are shown in Figure \ref{fig:dndF}. For demonstration, we consider $\alpha = 3.1$ (see the last section) and $\alpha = 1.5$ with $L_{\rm min } = 7000 {\rm W/Hz}$, $L_{\rm max} = 4.5\times 10^{13} {\rm W/Hz}$. 
The flux distribution from a spherical shell composes of three regimes: in order of increasing flux, 1) zero flux, 2) a power law which depends on the luminosity function, and 3) a power law with index -2.5.

@@

\section{Expected value for the highest flux from spherical shell at distance $r$}

Let us consider a flux distribution of the targets from spherical shell at distance $r$.

The probability of a star to have a flux between $F\sim F+dF$, $P(F)dF $, may be assumed to be proportional to the flux number distribution:
%%%
\begin{equation}
P(F) dF = \frac{dN_r}{dF} = C' F^{\alpha } r^{4-2\alpha }
\end{equation}
%%%
where
%%%
\begin{equation}
N_r = \frac{dN}{dr} 
\end{equation}
%%%
Note that $N_r$ is the {\it expected} number. \memoYF{not sure how to deal with the fact that this is expected number.}

Then, the probability to find one object with flux lower than $F'$ is
%%%
\begin{equation}
P (<F') = \int _{-\infty} ^{F'}  P (F) \; dF 
\end{equation}
%%%

The probability to find the maximum flux between $F_{\rm max} \sim F_{\rm max} + dF_{\rm max}$, $\mathcal{P} (F_{\rm max}) dF_{\rm max}$ is the joint probability where one of the targets has flux between $F_{\rm max} \sim F_{\rm max} + dF_{\rm max}$ and all others have smaller flux. Noting that we have $N_r$ ways to select the target with the highest flux, such a probability is:
%%%
\begin{equation}
\mathcal{P} (F_{\rm max}) dF_{\rm max} = N_R P (< F_{\rm max})^{N_R-1}  P(F_{\rm max}) dF_{\rm max} %\\
%&=& N_R P(F_{\rm max}) dF_{\rm max} \left[ 1 - P (> F_{\rm max}) \right]^{N_R-1} \\
%&\sim & N_R P(F_{\rm max}) dF_{\rm max} \left[ 1 - ( N_R -1 ) P (> F_{\rm max}) \right]
\end{equation}
%%%

The expected value for the maximum flux, $<F_{\rm max}>$, is obtained by calculating the flowing integral. 
%%%
\begin{eqnarray}
<F_{\rm max}> &=& \int dF_{\rm max} \, F_{\rm max} \mathcal{P} (F_{\rm max}) \\
&=& N_r \int dF_{\rm max} F_{\rm max} P (< F_{\rm max})^{N_r-1}  P(F_{\rm max})
\end{eqnarray}
%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{expectedFmax_alpha31.pdf}
   \plotoneh{expectedFmax_alpha15.pdf}
   \caption{Expected value for the maximum flux in $\mu {\rm Jy}$ from a spherical shell at distance $r$ with unit width $\Delta r = 1{\rm pc}$ (arbitrary) .}
  \label{fig:dnrdF}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Examples with $\alpha = 3.1$ and $\alpha = 1.5$ are shown in Figure \ref{fig:dnrdF}. For demonstration, we consider $\alpha = 3.1$ (see the last section) and $\alpha = 1.5$ with $L_{\rm min } = 7000 {\rm W/Hz}$, $L_{\rm max} = 4.5\times 10^{13} {\rm W/Hz}$. 
The flux distribution from a spherical shell composes of three regimes: in order of increasing flux, 1) zero flux, 2) a power law which depends on the luminosity function, and 3) a power law with index -2.5.

@@

In order to estimate the luminosity function of RGHJ emitters, one of the key factors is the distribution of mass-loss rates of giant host stars. 
For a given set of orbital parameters the luminosity ($L$) of a RGHJ depends on the stellar mass-loss rate and wind velocity as $\dot{M}_\star^{2/3} v^{5/3}$. 
The improved Reimers' equation \citep{reimers1975} given by \citet{schroder2005,schroder2007}:
%%%
\begin{equation}
\dot M_\star [M_\odot{\rm /yr}] \sim 8 \times 10^{-14} \frac{\tilde L \tilde R}{\tilde M} \left( \frac{T_{\rm eff}}{4000} \right)^{3.5} \left( 1 + \frac{g_{\odot }}{4300 g_{\star}} \right) \label{eq:mass-loss}
\end{equation}
%%%
implies $\dot{M}_\star \propto L_\star^{7.5/4} R_\star^{-0.75} M_\star^{-1} (1+A/g_\star)$ and crudely estimating the wind velocity as $v \propto \sqrt{GM/R} \propto M^{1/2} R^{-1/2}$, the luminosity of the RGHJ radio emission is given by $L \propto L_\star^{7/12} M_\star^{1/6} T_{\rm eff}^{8/3} (1+A/g_\star)^{2/3}$. 
Given that the dependences on $M_\star^{1/6}$ is relatively weak and that $T_{\rm eff}$ does not change significantly, the radio luminosity of RGHJs scales roughly as $L \propto L_{\star }^{7/12}$. 
Observations of globular clusters indicates the luminosity function of red giants is $dn/dL_{\star } \sim L_{\star }^{-1.8}$ (Sandquist et al. (1996)) %\citep{sandquist1996}
. Therefore, $\alpha \sim -3.1$.

The upper and lower limits are tentatively assumed as follows. 
%%%
\begin{itemize}
\item range of luminosity of ``evolved stars'':\\$L_{\odot }-10,000 L_{\odot }$
\item $\Rightarrow$ minimum radio luminosity of ``RGHJs''\\%
$L_{\rm min} = 2,1\times 10^{11}/30 {\rm MHz} = 7000 {\rm W/Hz}$
\item $\Rightarrow $maximum radio luminosity of ``RGHJs'':\\$L_{\rm max} = 4.5\times 10^{13} {\rm W}$
\end{itemize}
%%%

>>>>

\title{Luminosity function and flux function}
\author{%
Yuka Fujii\altaffilmark{1}, %
David S. Spiegel\altaffilmark{2,3,4}%
}

@@

\begin{abstract}
We examine how the highest observed flux evolves as we observe further. 
This question is motivated because it appears to be a competition between the fact that the observed flux decreases as the objects are located further, and the fact that the number of targets increases as we observe further. 
%
To answer this question, we calculate the expected value of the highest flux from a given spherical shell and see whether the expected value increases as a function of distance (radius) of the sphere. 
In doing so, we consider  a luminosity function characterized by a power law with lower and upper limits of luminosity, %
We find the expected value of the highest flux monotonically decreases as a function of radius of the spherical shell, and thus we are not likely to find the brightest object far away. 
%we calculate the expected value of the highest flux from a given spherical shell and see whether the expected value increases as a function of distance (radius) of the sphere. 
%firstly we consider the distribution of observed flux from a sphere within distance $R$ of a given luminosity function ("flux function"). The flux function composes of three regimes: in order of increasing flux, 1) zero flux, 2) a power law which depends on the luminosity function, and 3) a power law with index -2.5. 
%Then, we compute the expected value of the highest flux from a spherical shell at distance $R\sim R+dR$. within a volume with radius $R$. 
\end{abstract}

%\keywords{planetary systems --- planets and satellites: general}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Distribution of observed flux from a spherical shell} 
\label{sec:luminosity function}

@@

In this section, we consider the number density of observed flux of the targets at spherical shell at distance $r$. 
In the following, Luminosity and (observed) flux are denoted by $L$ and $F$, respectively. 
\vspace{1\baselineskip}

@@

Now, we want to change variables from $\{L,\,r\}$ to $\{F,\,r\}$. 
The relationship between flux $F$ and luminosity $L$ is, as usual,
%%%
\begin{equation}
F = \frac{L}{4\pi r^2} 
\end{equation}
%%%
where $r$ is the distance to the target. 
As changing the set of variables needs to employ Jacobian,
%%%
\begin{eqnarray}
\frac{d^2 N}{dL dr} &=& \Psi (4\pi r^2F', r)\; det(J) \\
det(J) &=& \left| \PD{r}{r}\PD{L}{F'} - \PD{r}{F'}\PD{L}{r} \right| = 4 \pi r^2
\end{eqnarray}
%%%
Therefore, the number density of the flux coming from the spherical shell is
%%%
\begin{equation}
\frac{d^2 N}{dF dr} = \left\{
\begin{array}{ll}
C' F^{-\alpha} r^{4-2\alpha } & (L_{\rm min} < 4\pi r^2 F < L_{\rm max}) \\
0 & (\mbox{elsewhere})
\end{array}
\right. \label{eq:dNdFdr}
\end{equation}
%%%
where $C' \equiv (4 \pi )^{2-\alpha}$.

\section{Distribution of observed flux within a sphere}

@@

\vspace{1\baselineskip}

Examples with $\alpha = 3.1$ and $\alpha = 1.5$ are shown in Figure \ref{fig:dndF}. For demonstration, we consider $L_{\rm min } = 7\times 10^3~{\rm W/Hz}$, $L_{\rm max} = 1.5\times 10^6~{\rm W/Hz}$ for some reasons (Appendix). 
The flux distribution from a sphere composes of three regimes: in the order of increasing flux, 1) zero flux, 2) a power law which depends on the luminosity function, and 3) a power law with index -2.5.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{dndF_alpha31.pdf}
   \plotoneh{dndF_alpha15.pdf}
   \caption{Flux distribution of targets within radius $R$ with varying power-law index and $L_{\rm min } = 7\times 10^3~{\rm W/Hz}$, $L_{\rm max} = 1.5\times 10^6~{\rm W/Hz}$. The $y$-axis is arbitrary. }
  \label{fig:dndF}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Expected value for the highest flux from a spherical shell at distance $r$}

Let us consider a flux distribution of the targets from a spherical shell at distance $r$ with thickness $dr$. 
The ({\it expected}) total number of the target within that spherical shell is
%%%
\begin{equation}
4 \pi r^2 dr \int _0 ^{\infty} \frac{d^2 N}{dF~dr} dF \equiv n_r. 
\end{equation}
%%%
\memoYF{not sure how to deal with the fact that this is expected number.}

The probability for a star within the spherical shell to have a flux between $F\sim F+dF$, $P(F)dF $, may be assumed to be proportional to the flux number distribution given by equation (\ref{eq:dNdFdr}):
%%%
\begin{equation}
P(F) dF = C' F^{\alpha } r^{4-2\alpha }
\end{equation}
%%%
where
%%%
\begin{equation}
C' = n_r \left[ \int _0 ^{\infty}P(F)~dF~\right]^{-1}
\end{equation}
%%%

The probability to find the maximum flux between $F_{\rm max} \sim F_{\rm max} + dF_{\rm max}$, $\mathcal{P} (F_{\rm max}) dF_{\rm max}$ is the joint probability where one of the targets has flux between $F_{\rm max} \sim F_{\rm max} + dF_{\rm max}$ and all others have smaller flux. 
The probability to find one object with flux lower than $F'$ is
%%%
\begin{equation}
P (<F') = \int _{-\infty} ^{F'}  P (F) \; dF 
\end{equation}
%%%
Noting that we have $n_r$ ways to select the target with the highest flux, such a probability is:
%%%
\begin{equation}
\mathcal{P} (F_{\rm max}) dF_{\rm max} = n_r P (< F_{\rm max})^{n_r-1}  P(F_{\rm max}) dF_{\rm max} %\\
%&=& N_R P(F_{\rm max}) dF_{\rm max} \left[ 1 - P (> F_{\rm max}) \right]^{N_R-1} \\
%&\sim & N_R P(F_{\rm max}) dF_{\rm max} \left[ 1 - ( N_R -1 ) P (> F_{\rm max}) \right]
\end{equation}
%%%

The expected value for the maximum flux among the targets within a spherical shell at radius $r$ (with width $dr$), $<F_{\rm max}>$, is obtained by calculating the flowing integral. 
%%%
\begin{eqnarray}
<F_{\rm max}> &=& \int dF_{\rm max} \, F_{\rm max} \mathcal{P} (F_{\rm max}) \\
&=& n_r \int dF_{\rm max} F_{\rm max} P (< F_{\rm max})^{n_r-1}  P(F_{\rm max})
\end{eqnarray}
%%%

Figure \ref{fig:expectedFmax} shows $<F_{\rm max}>$ as a function of distance $r$ with varying number density of the targets. 
Again, examples for $\alpha = 3.1$ and $\alpha = 1.5$, with $L_{\rm min } = 7.0\times 10^3 ~{\rm W/Hz}$ and $L_{\rm max} = 1.5\times 10^{13} ~{\rm W/Hz}$ are shown. 
It is seen that the expected values monotonically decrease as we observe further. This indicates the brightest object is the closest one. \memoYF{??}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htbp]
   \plotoneh{expectedFmax_alpha31.pdf}
   \plotoneh{expectedFmax_alpha15.pdf}
   \caption{Expected value for the maximum flux in $\mu {\rm Jy}$ from a spherical shell at distance $r$ with unit width $\Delta r = 1{\rm pc}$ (arbitrary) .}
  \label{fig:expectedFmax}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In order to estimate the luminosity function of RGHJ emitters, one of the key factors is the distribution of mass-loss rates of giant host stars. 
For a given set of orbital parameters the luminosity ($L$) of a RGHJ depends on the stellar mass-loss rate and wind velocity as $\dot{M}_\star^{2/3} v^{5/3}$. 
The improved Reimers' equation \citep{reimers1975} given by \citet{schroder2005,schroder2007} says:
%%%
\begin{equation}
\dot M_\star [M_\odot{\rm /yr}] \sim 8 \times 10^{-14} \frac{\tilde L \tilde R}{\tilde M} \left( \frac{T_{\rm eff}}{4000} \right)^{3.5} \left( 1 + \frac{g_{\odot }}{4300 g_{\star}} \right), \label{eq:mass-loss}
\end{equation}
%%%
which implies $\dot{M}_\star \propto L_\star^{7.5/4} R_\star^{-0.75} M_\star^{-1} (1+A/g_\star)$. 
By crudely estimating the wind velocity as $v \propto \sqrt{GM/R} \propto M^{1/2} R^{-1/2}$, the luminosity of the RGHJ radio emission is given by $L \propto L_\star^{7/12} M_\star^{1/6} T_{\rm eff}^{8/3} (1+A/g_\star)^{2/3}$. 
Given that the dependences on $M_\star^{1/6}$ is relatively weak and that $T_{\rm eff}$ does not change significantly, the radio luminosity of RGHJs scales roughly as $L \propto L_{\star }^{7/12}$. 
Observations of globular clusters indicates the luminosity function of red giants is $dn/dL_{\star } \sim L_{\star }^{-1.8}$ (Sandquist et al. (1996)) %\citep{sandquist1996}
. Therefore, $\alpha \sim -3.1$.

The upper and lower limits are tentatively assumed as follows. 
%%%
\begin{itemize}
\item range of luminosity of ``evolved stars'':\\$L_{\odot }-10,000 L_{\odot }$
\item $\Rightarrow$ minimum radio luminosity of ``RGHJs''\\%
$L_{\rm min} = 2.1\times 10^{11} W/30\,{\rm MHz} = 7.0 \times 10^3\,{\rm W/Hz}$
\item $\Rightarrow $maximum radio luminosity of ``RGHJs'':\\$L_{\rm max} = 4.5\times 10^{13}{\rm W}/30\,{\rm MHz} = 1.5 \times 10^{6}\,{\rm W/Hz}$
\end{itemize}
%%%
